---
title: "R Davis class lecture Week 2"
output: html_document
---

#how to use and manipulate vectors
#indexing and subetting are the same thing!

```{r}
letters
a<- 1
a<- 2
```
```{r}
gracevec<- c(7,8,9)
gracevec
```

```{r}
gracevec[2]
```


```{r}
gracevec[c(2,2)]
```

```{r}
gracevec[c(1,2,3)]
gracevec[c(3,2,1)]
```

# strings are the things that you put in quotes, like column names etc. A string is a character vector or something 

#can use the ? to find out more about certain functions 


```{r}
(gracestring <- c("7", "17", "77","b"))
gracestring
```

```{r}
data.frame(c(gracestring,gracestring))
```

#can create a dataframe in R by using the setting 



```{r}
data.frame(first_column = gstring,second_column = c(gstring,'23'))

```

#you can use class to figure out what a column or row contains (i.e. numbers vs characters)


```{r}
class(gracestring)
```


# You can use a list to store random things in R

```{r}
test_list = list("a","b","c","d")
test_list

```



```{r}
c(test_list,2)
str(test_list)
test_list[[1]][2] <- 'add next thing'


```
```{r}

gracelist <- list(gracestring,"a","c")
gracelist
```

```{r}
vectorgracelist <- c("a","c")
vectorgracelist
```

