---
title: "R-Davis_Week9hw"
output: html_document
---


Use the README file associated with the Mauna Loa dataset to determine in what time zone the data are reported, and how missing values are reported in each column. With the mloa data.frame, remove observations with missing values in rel_humid, temp_C_2m, and windSpeed_m_s. 

Generate a column called “datetime” using the year, month, day, hour24, and min columns. Next, create a column called “datetimeLocal” that converts the datetime column to Pacific/Honolulu time (HINT: look at the lubridate function called with_tz()). 

Then, use dplyr to calculate the mean hourly temperature each month using the temp_C_2m column and the datetimeLocal columns. (HINT: Look at the lubridate functions called month() and hour()).

Finally, make a ggplot scatterplot of the mean monthly temperature, with points colored by local hour.

```{r}
library(lubridate)
library(tidyverse)
```


```{r}
loa <- read.csv("https://raw.githubusercontent.com/gge-ucd/R-DAVIS/master/data/mauna_loa_met_2001_minute.csv")
```

```{r}
loa
```
```{r}
nona <- loa %>%
  filter(rel_humidn = -99, TRUE)

#original attempt! was trying to get it to tell me which values were false or true for -99
```

```{r}
mloa_sub = loa %>% filter(rel_humid!=-99,temp_C_2m!=-999.9,windSpeed_m_s!=-99.9)


```

```{r}
#Generate a column called “datetime” using the year, month, day, hour24, and min columns. Next, create a column called “datetimeLocal” that converts the datetime column to Pacific/Honolulu time (HINT: look at the lubridate function called with_tz()). Then, use dplyr to calculate the mean hourly temperature each month using the temp_C_2m column and the datetimeLocal columns. (HINT: Look at the lubridate functions called month() and hour()). Finally, make a ggplot scatterplot of the mean monthly temperature, with points colored by local hour.
# this version pastes year month day together by row and hten calls ymd
mloa_sub %>% mutate(datetime = paste(year,month,day,sep = '-')) %>%
  mutate(datetime = ymd(datetime)) %>% select(datetime)
#this version does both actions at once in the same mutate call
mloa_sub %>% mutate(datetime = ymd(paste(year,month,day,sep = '-'))) %>% select(datetime)

#now let's add in hour24...
mloa_sub %>% mutate(datetime = ymd_h(paste(year,month,day,hour24,sep = '-'))) %>% select(datetime)
```

```{r}
?with_tz()
```

x <- ymd_hms("2009-08-07 00:00:01", tz = "America/New_York")
with_tz(x, "GMT")

```{r}

#datetimeLoc <- mloa_sub(mutate(datetimeLocal = with_tz(datetime, tzone = "PST"))

#attempt #1
```


```{r}
mloa_sub %>% mutate(datetime = ymd_h(paste(year,month,day,hour24,sep = '-'))) %>% select(datetime)
mloa_sub = mloa_sub %>%
 mutate(datetime = ymd_hm(paste(year,month,day,hour24,min)))

```


```{r}
mloa_sub$datetime = with_tz(mloa_sub$datetime, tz= "Pacific/Honolulu")                     
mloa_sub$datetime

```

```{r}
mloa_sub
```

Then, use dplyr to calculate the mean hourly temperature each month using the temp_C_2m column and the datetimeLocal columns. (HINT: Look at the lubridate functions called month() and hour()).

```{r}
library(dplyr)
```

month(x, label = FALSE, abbr = TRUE, locale = Sys.getlocale("LC_TIME"))

month(x) <- value

```{r}
?lubridate::month

```

```{r}

mloa_sub$month <- mean(mloa_sub$temp_C_2m)


#attempt 1!
```



Challenge
This challenge will deal with countries’ population growth. To access the data, load (and install, if needed) the gapminder library and access its life expectancy dataset using:
```{r}
#install.packages("gapminder")
```

```{r}

library(gapminder)
d <- gapminder::gapminder
```

Write a new function that takes two arguments, the gapminder data.frame (d) and the name of a country (e.g. "Afghanistan"), and plots a time series of the country’s population. That is, the return value from the function should be a ggplot object. - It is often easier to modify existing code than to start from scratch. To start out with, try writing some code that works interactively, and then wrap it in a function.

myfunction <- function(x) {
  top <- sum(is.na(x))
  bottom <- sum(!is.na(x))
  prop <- top/bottom
}


```{r}
tab <- table(gapminder$country)
tab
```

```{r}
myfunction <- function(x) {
  gap<- gapminder$pop
  country<- gapminder$year
  plot <- gap*country
}

```















