---
title: "R-Davis_Week5hw"
output: html_document
---

This week’s questions will have us practicing pivots and conditional statements.

  




Create a tibble named surveys from the portal_data_joined.csv file. 

```{r}
library(tidyverse)
surveys <- read_csv("data/portal_data_joined.csv")
```


```{r}
surveys
```

Then manipulate surveys to create a new dataframe called surveys_wide with a column for genus and a column named after every plot type, with each of these columns containing the mean hindfoot length of animals in that plot type and genus. So every row has a genus and then a mean hindfoot length value for every plot type.The dataframe should be sorted by values in the Control plot type column. This question will involve quite a few of the functions you’ve used so far, and it may be useful to sketch out the steps to get to the final result.

# I understood some of this question, but got lost on how to combine all of the parts together correctly. 
```{r}
surveys_wide <- surveys %>%
  filter(is.na(hindfoot_length)) %>%
  group_by(genus, plot_type) %>%
  summarise(mean_hindfoot = mean(hindfoot_length)) %>% 
  pivot_wider(names_from = 'plot_type',values_from = "mean_hindfoot" ) %>%
  arrange(Control) # im not really sure what arrange control does 
surveys_wide
```


Using the original surveys dataframe, use the two different functions we laid out for conditional statements, ifelse() and case_when(), to calculate a new weight category variable called weight_cat. For this variable, define the rodent weight into three categories, where “small” is less than or equal to the 1st quartile of weight distribution, “medium” is between (but not inclusive) the 1st and 3rd quartile, and “large” is any weight greater than or equal to the 3rd quartile. (Hint: the summary() function on a column summarizes the distribution). For ifelse() and case_when(), compare what happens to the weight values of NA, depending on how you specify your arguments.

```{r}
#start by making rodent weight into 3 categories

#need to filter surveys to just be rodent taxa 

catsurveys <- surveys %>%
  group_by(taxa, weight) 
  
catsurveys

#I am unsure how to select for just rodent taxa so that I can get the quartile values using the summary function on the rodent taxa only 

summary(surveys$taxa, "Rodent")

#that was my best guess 
```


Here is the given answer 

```{r}
summary(surveys$weight)

```

lets try again 

```{r}

catsurveys <- surveys %>%
  group_by(taxa, weight) 
  if_else(surveys$weight <= 20, "small", "big")
  
  #could not figure out how to add a small, big, and a medium category 
```

Given answer 

```{r}
surveys %>% 
  mutate(weight_cat = case_when(
    weight >= 20.00 ~ "small",
    weight > 20.00 & weight < 48.00 ~ "medium",
    T ~ "large"
  ))

#looks like if I used an "&" I can add additional weight categories 
```

lets try again

```{r}
atsurveys <- surveys %>%
  group_by(taxa, weight) 
  mutate(weight_cat = case_when(surveys$weight <= 20 ~"small", surveys$weight >20 & surveys$weight < 48 ~"medium", surveys$weight >48 ~ "big"))
  
#weight_cat is not an already existing variable, but a new variable you are creating with the "mutate" function 
  

#R did not like this way either :( 
```


```{r}
surveys %>% 
  mutate(weight_cat = ifelse(weight >= 20.00, "small",
                             ifelse(weight > 20.00 & weight < 48.00, "medium","large")))
```










