---
title: "R-Davis_Week6hw"
output: html_document
---

For our week six homework, we are going to be practicing the skills we learned with ggplot during class. You will be happy to know that we are going to be using a brand new data set called gapminder. This data set is looking at statistics for a few different country's including population, GDP per capita, and life expectancy. Download the data using the code below. Remember, this code is looking for a folder called data to put the .csv in, so make sure you have a folder named data, or modify the code to the correct folder name.

```{r}
library(tidyverse)

gapminder <- read_csv("https://gge-ucd.github.io/R-DAVIS/data/gapminder.csv") #ONLY change the "data" part of this path if necessary
```

First calculate mean life expectancy on each continent. Then create a plot that shows how life expectancy has changed over time in each continent. Try to do this all in one step using pipes! (aka, try not to create intermediate dataframes)



```{r}
gapminder
```
```{r}
mean <- gapminder %>%
  filter(!is.na(lifeExp)) %>%
  group_by(continent, lifeExp) %>%
  summarise(meanlife, mean(lifeExp))
mean
#summarise function allows you to calculate summary statistics

```


```{r}

#now lets add that plot! shows how life expectancy has changed over time in each continent
timeplot <- gapminder %>%
  filter(!is.na(year))
  group_by(lifeExp,year,continent) %>%
    ggplot(aes(x= year, y=lifeExp))

#it cannot find lifeExp... not quite sure what I did wrong there 
  #tried adding quotes around the words, that was not the way 
  
```
```{r}
timeplot <- gapminder %>%
  group_by(lifeExp,year,continent) %>%
  ggplot(aes(x = year, y = lifeExp, color = continent))
    
timeplot

#okay, now it can find the "LifeExp" once I removed the filter column, coolio I guess
#well shucks, this does not look like a plot of anything. Back to the drawing board 


```
```{r}
timeplot2.0 <- gapminder %>%
  group_by(lifeExp,year,continent) %>%
 ggplot()+
  geom_point(aes(x = year, y = lifeExp, color = continent))
timeplot2.0

#welp that's a somewhat pretty and understandable plot 
```
```{r}
timeplot3.0 <- gapminder %>%
  group_by(lifeExp,year,continent) %>%
 ggplot()+
  geom_point(aes(x = year, y = mean(lifeExp), color = continent))
timeplot3.0

#well thats not right either 
```


Onwards! 


Look at the following code and answer the following questions. What do you think the scale_x_log10() line of code is achieving? What about the geom_smooth() line of code?



```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) +
    geom_point(aes(color = continent), size = .25) + 
    scale_x_log10() +
    geom_smooth(method = 'lm', color = 'black', linetype = 'dashed') +
    theme_bw()

#well, I think the best way to figure out what these lines of codes do is to remove them! 

```


```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) +
    geom_point(aes(color = continent), size = .25) + 
    geom_smooth(method = 'lm', color = 'black', linetype = 'dashed') +
    theme_bw()

#okay, so here we can see that the scale_X_log10 code scales the x axis in such as way that allows the plot to be more easily read and interpreted 
```
```{r}
ggplot(gapminder, aes(x = gdpPercap, y = lifeExp)) +
    geom_point(aes(color = continent), size = .25) + 
    scale_x_log10() +
    theme_bw()

#so here it looks like geom_smooth adds a line that best fits the overall trend of the data! Coolio! 
```


Create a boxplot that shows the life expectency for Brazil, China, El Salvador, Niger, and the United States, with the data points in the backgroud using geom_jitter. Label the X and Y axis with “Country” and “Life Expectancy” and title the plot “Life Expectancy of Five Countries”.


```{r}
gapminder
```

```{r}
boxbud <- gapminder %>%
  filter(country = "Brazil","China","El Salvador","Niger","United States")

#nope 
```

```{r}
newgap <- c(gapminder$country, "Brazil", "Niger")
newgap

#also no 
```


```{r}

#Okay this here makes some sense theoretically, but I would have never arrived here without peeking at the answers 

countries <- c("Brazil", "China", "El Salvador", "Niger", "United States") #create a vector with just the countries we are interested in

gapminder %>% 
  filter(country %in% countries) %>% #will watch the video that goes over this %in% thing 
  ggplot(aes(x = country, y = lifeExp))+ 
  geom_boxplot() +
  geom_jitter(alpha = 0.3, color = "blue")+
  theme_minimal() +
  ggtitle("Life Expectancy of Five Countries") + #title the figure
  theme(plot.title = element_text(hjust = 0.5)) + #centered the plot title
  xlab("Country") + ylab("Life Expectancy") #how to change axis names

# the amount we add plots to plots is also quite confusing to me 
```














