---
title: "R-Davis_Week3Homework-grosburg"
output: html_document
---


#'head' function allows you to see only the first 6 rows of all your columns or something

#Can use the 'identical' function to ask if two different data subsets you made in R are the same  

#Can use the mutate function to add a new column to a dataframe that is based on a existing column you manipulated 


```{r}
library(tidyverse)
```

#Load your survey data frame with the read.csv() function. 


```{r}
surveys <- read.csv('data/portal_data_joined.csv')
surveys
```
#Create a new data frame called surveys_base with only the species_id, the weight, and the plot_type columns.

```{r}
surveys_base <- select(surveys, "species_id","weight","plot_type")
str(surveys_base)
```

# Have this data frame only be the first 60 rows. 


```{r}
surveys_base <- surveys_base[1:60, c(1:3)]
str(surveys_base)
```


#Convert both species_id and plot_type to factors. Explore these variables and try to explain why a factor is different from a character.


```{r}
surveys_base$species_id <- as.factor(surveys_base$species_id)
class(surveys_base$species_id)

```


```{r}
surveys_base$plot_type <- as.factor(surveys_base$plot_type)
class(surveys_base$plot_type)
```

#you can use the class function to tell you wether a column is a factor or a character

#A factor can be treated like a number whereas a character is treated more like a name of something or a descriptor of something and not something you can do numerical stuff with as much
`


#Remove all rows where there is an NA in the weight column.

```{r}
?na.omit
```



```{r}
is.na(surveys_base$weight)
```
```{r}
!is.na(surveys_base$weight)
```


```{r}
surveys_base[is.na(surveys_base$weight),]
```
```{r}
surveys_base[!is.na(surveys_base$weight),]
```
```{r}
surveys_base[complete.cases(surveys_base),]
```
```{r}
complete.cases(surveys_base)
```




```{r}
surveys_base %>% na.omit(surveys_base$weight)
```

#Create a new data frame from the surveys that meets the following criteria
#contains only the species_id coulmn 
#and a new column called hondfoot_half containing values that are half the hindfoot_half 
#in hindfoot_half, make it where there are no Nas, and avlues are less than 30 

```{r}
surveys$hindfoot_length

surveys_hindfoot_half <- surveys %>%
  filter(.,!is.na(hindfoot_length))%>%
  mutate(.,hindfoot_half = hindfoot_length/2) %>%
  select(.,species_id,hindfoot_half) %>%
  filter(.,hindfoot_half < 30)

str(surveys_hindfoot_half)
```


```{r}

surveys_hindfoot_half %>% group_by(species_id) %>%
  mutate(hindfoot_length = hindfoot_half*2) %>%
  summarize(mean(hindfoot_length),min(hindfoot_length,max(hindfoot_length))) %>%
  head()
```








